---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---

###################################### 
## 1. RUN: LOAD PACKAGES
```{r Setup}
library(readxl) # Read excel
library(cmdstanr) # Interface to Stan
library(posterior) # Process Stan draws
library(doParallel); library(parallel) # R Multi-threading for EB
library(shiny)
source("https://raw.githubusercontent.com/klattery/Estimation-Ecosystem/master/EE_Functions_Stan4.0.R")
source("https://raw.githubusercontent.com/klattery/Estimation-Ecosystem/master/Stan_RShiny_UI.R")
filename <- list() # Placeholder for filenames to import (data, covariates, excel specs)
#set_cmdstan_path("/home/rstudio/cmdstan")
dir <- list(data = "/home/rstudio/CBC/data", # data files
            work = "C:/Users/K.Lattery/OneDrive - SKIM/Temp", # output and other work
            stanmodel = "/home/rstudio/CBC/StanCode", # code for Stan Model
            stanout = "C:/Users/K.Lattery/OneDrive - SKIM/Temp" # Stan output 
            )
```

############################################### 
## *2. FORM POPS-UP TO SPECIFY FILES*
#### `MANDATORY INPUT FORM`
#### VERIFY data_conjoint, specs\_, data_cov
```{r Load Data}
shinyApp(env_shiny$ui_1, env_shiny$server_1)
cat("Check Step 2 imported your data accurately:
  data_conjoint = stacked data file
  specs_*       = Excel specs")
```

################################################ 
## 3. RUN: CODE AND PREPARE DATA FOR STAN
#### CHECK: PRINTED OUTPUT LOOKS RIGHT
```{r Code and Prepare}
col_id_task_dep <- c(1,2,ncol(data_conjoint)) # Columns for id, task, dep
# 3) Code and Prepare data_stan =============
indcode_spec <- indcode_spec_files(data_conjoint, specs_att_coding,
                                   specs_pair_constraints, 
                                   add_none, col_id_task_dep[3]) # Code and constraints each attribute
# Here is where you would do any custom coding indcode_spec[[i]] <- "syntax"
if (!is.null(indcode_spec)){
  indcode_list <- make_codefiles(indcode_spec) # Combine specifications above into one list
  save_codemastercon(indcode_list,dir$work, out_prefix) # Save combined code_master and constraints
  data_stan <- prep_file_stan(idtaskdep = data_conjoint[,col_id_task_dep],
                              indcode_list = indcode_list,
                              data_cov = data_cov, specs_cov_coding = specs_cov_coding,
                              check_collinearity = check_collinearity)
  if (add_none){
    if  (is.null(indcode_spec$None$message)){
      cat("\nDid not Add a 'None' as no rows of all 0 found.\n")
    } else cat(indcode_spec$None$message)  
  } 
} else cat("!! STOP NOW and FIX NAMES that do not match !!")

```


#################################################### 
## 4. OPTIONAL: Estimate Aggregate Model
#### Chunk runs only if est_aggmodel is TRUE
```{r Aggregate Model}
if (est_aggmodel){
agg_beta <- est_agg_model(data_stan, maxit = 200, reltol = 1e-5)
agg_beta_r <- data_stan$code_master %*% agg_beta$par
write.table(agg_beta_r, file = file.path(dir$work, paste0(out_prefix,"_agg_beta.csv")), sep = ",", na = ".", row.names = TRUE)
cat("\nAggregate model results in R file: agg_beta_r")
} else cat("Aggregate model not estimated")
```

################################################ 
## 4a. Generate sim_pop and predictions for sim_pop
```{r}
if (exists("agg_beta")){
  alpha <- agg_beta$par
} else alpha <- rnorm(data_stan$P, 0, 2) # Customize if desired
sim_pop <- gen_sim_pop(alpha, 1.5 * data_stan$prior_cov, data_stan$I,
                       draw_alpha = FALSE, draw_cov = TRUE,
                       con_sign = data_stan$con_sign)
data_stan$sim_dep <- sim_MNL(data_stan, sim_pop)
message("Hit Rate of Choice w/Error = ",
        sum(data_stan$sim_dep$choice_noerr * data_stan$sim_dep$choice_werr)/sum(data_stan$sim_dep$choice_werr))

```

################################################### 
## 5. OPTIONALLY CHANGE DEFAULTS
#### THEN RUN TO ESTIMATE MODEL
```{r HB Estimate}
# 5) Modeling Options. Defaults are usually fine 
data_model <- list(
  dep = data_stan$sim_dep$choice_werr,
  iter_warmup = 400, # warmup of 400 is plenty
  iter_sampling = 400, # sampling of 400 is plenty
  df = 2,              # recommend df = 2 for Wishart
  prior_cov_scale = c(lb = .8, ub = 2,
                      mean = 1, sigma = 2), # prior_sd ~ normal(mean, sigma)[lb, ub]
  con_use = 1, # 0 = ignore constraints, 1 = use
  con_factors = c(mult = .25, bound = 1.5), # Default (.25, 1.5) must be >0
  threads_per_chain = data_stan$threads_rec,
  agg_model = NULL, tag = NULL, ind = NULL, resp_id = NULL
)
data_model$splitsize <- round(.5 + data_stan$T/(4 * data_model$threads_per_chain))
stan_file <- "BaseHB_wPairCon_v2.5_cholscalediag.stan" # Name of stan model in dir$stanmodel
##################################
# ESTIMATE MODEL =============

stan_outname <- paste0(out_prefix, "_StanOut_", 
                  format(Sys.time(), '%Y%m%d-%H%M%S')) # Base Name of Stan Output files  
rm(indcode_spec); rm(indcode_list); gc() # Clear RAM
#####  Compile and RunStan Model 
HB_model <- cmdstan_model(file.path(dir$stanmodel,stan_file), quiet = TRUE,
                          cpp_options = list(stan_threads = TRUE))
message_estimation(dir, stan_outname)
HB_fit <- HB_model$sample(modifyList(data_stan, data_model),
                iter_warmup = data_model$iter_warmup,
                iter_sampling = data_model$iter_sampling,
                output_dir = dir$stanout,
                output_basename = stan_outname, # set stan_outname above if changing
                chains = 2,
                parallel_chains = 2,
                threads_per_chain = data_model$threads_per_chain,
                save_warmup = TRUE,
                refresh = 10,
                adapt_delta = .8,
                seed = 123,
                init = .1,
                show_messages = FALSE,
                diagnostics = NULL)
saveRDS(HB_fit, file.path(dir$work, "HB_fit.rds"))
#####  Check Convergence, Export Files  
if (min(HB_fit$return_codes() == 0)){
  cat("Reading draws from Stan csv output into R (large files take time)...")
  draws_beta <- read_cmdstan_csv(HB_fit$output_files(), variables = "beta_ind", format = "draws_list")
  draws_beta$warmup_draws <- NULL # to save space
  utilities <- get_mean_beta(draws_beta) # computes mean point estimates from draws
  checkconverge_export(draws_beta, vnames = colnames(data_stan$code_master), out_prefix, dir$work)
  process_utilities(data_stan, utilities, out_prefix, dir$work)
  if (data_stan$P_cov > 0) covariate_means <- get_covariates(HB_fit, data_stan)
} else message("Stan Estimation Did not Finish")
##################################
```

## OPTIONAL EMPIRICAL BAYES
```{r EB}
# 6) Optional Empirical Bayes =============
linux <- (Sys.info()[1] != "Windows")
if (linux) r_cores <- min(8,max(detectCores() -1,1)) # Number of cores to use for EB
if (!linux) r_cores <- min(4,max(detectCores() -1,1)) # Number of cores to use for EB
eb_betas_est(data_stan, draws_beta, data_stan$x0, r_cores,
               out_prefix, dir$work, cov_scale = 1, linux)
```
